Elastic Scattering in General Relativistic Ray Tracing for Neutrinos

### 2016-11-29 paper.tex
Add some section labels and other odds and ends
Move a figure to hack fix the float error
Compiling the paper at the previous commit produces a seemingly
unrelated "undefined control sequence" error at a \todo statement
following the figure. If I insert another \todo statement right
after the figure, the error shifts to earlier in the doc by one
\todo statement. Turning off the \todos (\renewcommand) removes
the errror, but in the meantime, I'll simply move the figure till
it falls on another page. That also removes the error.

Some possibly informative stackexchange dialogs:
http://tex.stacke

\subsection{The Rendering Equation}
\label{ssec:rendering_eqn}
Defining the source function
$\mathscr{S}^*\equiv\mathscr{E}^*/\mathscr{K}^*$,
we can integrate Eqn.~\ref{eqn:boltzmann_linear} directly,
\begin{equation}
  \label{eqn:rendering}
  f(\lambda) = f(\lambda_e) e^{-\tau(\lambda,\lambda_e)} +
  \int_{\lambda_e}^\lambda d\lambda'\,e^{-\tau(\lambda,\lambda')}
  \mathscr{K}^*(\lambda') \mathscr{S}^*(\lambda'),
\end{equation}
where the optical depth is defined,
\begin{equation}
  \label{eqn:optical_depth}
  \tau(\lambda,\lambda') \equiv -\int_{\lambda'}^\lambda
  d\lambda'' \, \mathscr{K}^*(\lambda''),
\end{equation}
and the parameterization conventions are depicted in
Fig.~\ref{fig:affine_param}.

@@

\subsection{Numerical Implementation}
\label{ssec:numerical}
Much of our numerical implementation is borrowed from the geodesic evolution
system described in~\cite{bohn2016-code}.
We integrate Eqns.~\ref{eqn:geodesic_x} and~\ref{eqn:geodesic_p} in the form given by
\citep{hugh1994-eh_finding}.
\todo{describe eqns for $f_{\rm integ}$ and $\tau$}
\todo{describe $\nu$-specific numerics}

>>>>

\subsection{The Rendering Equation}
\label{ssec:rendering_eqn}
Defining the source function
$\mathscr{S}^*\equiv\mathscr{E}^*/\mathscr{K}^*$,
we can integrate Eqn.~\ref{eqn:boltzmann_linear} directly,
\begin{align}
  \label{eqn:rendering}
  f(\lambda,\lambda_e) &=
  f_{\rm bound}(\lambda,\lambda_e) + f_{\rm integ}(\lambda,\lambda_e), \\
  \label{eqn:rendering_fbound}
  f_{\rm bound}(\lambda,\lambda_e) &=
  f(\lambda_e) e^{-\tau(\lambda,\lambda_e)}, \\
  \label{eqn:rendering_finteg}
  f_{\rm integ}(\lambda,\lambda_e) &=
  \int_{\lambda_e}^\lambda d\lambda'\,e^{-\tau(\lambda,\lambda')}
  \mathscr{K}^*(\lambda') \mathscr{S}^*(\lambda'),
\end{align}
where the optical depth is defined,
\todo{rewrite this section simpler; don't give integral solution
  since we don't use}
\begin{equation}
  \label{eqn:optical_depth}
  \tau(\lambda,\lambda') \equiv -\int_{\lambda'}^\lambda
  d\lambda'' \, \mathscr{K}^*(\lambda''),
\end{equation}
and the parameterization conventions are depicted in
Fig.~\ref{fig:affine_param}.

@@

\subsection{Numerical Implementation}
\label{ssec:numerical}
Much of our numerical implementation is borrowed from the geodesic evolution
system described in~\cite{bohn2016-code}.
We integrate Eqns.~\ref{eqn:geodesic_x} and~\ref{eqn:geodesic_p} in the form
given by \citep{hugh1994-eh_finding},
and Eqns.~\ref{eqn:rendering} and~\ref{eqn:optical_depth}
in the form given below.
By using the time-component of Eqn.~\ref{eqn:geodesic_x} ($dt=d\lambda \, p^t$)
we may transform our integrations to coordinate time.
The equations are
\begin{align}
  \label{eqn:num_x}
  \frac{dx^i}{dt} &=
  g^{ij} \frac{p_j}{p^t} - \beta^{i}\\
  \label{eqn:num_p}
  \frac{dp_i}{dt} &=
  -\alpha \alpha_{,i} p^t
  + \beta^k_{,i} p_k
  - \frac{1}{2} g^{jk}_{,i}\frac{p_j p_k}{p^t}\\
  \label{eqn:num_finteg}
  \frac{df_{\rm integ}}{dt} &=
  e^{-\tau} \mathscr{K}^* \mathscr{S}^*\\
  \label{eqn:num_tau}
  \frac{d\tau}{dt} &=
  -\mathscr{K}^*
\end{align}
where the comma denotes a partial derivative
and we have used the standard 3+1 decomposition of the metric
\begin{equation}
  \label{eqn:adm_metric}
  \psi_{\mu\gamma} :=
  \left(
  \begin{matrix}
    -\alpha^2 + \beta^i \beta_i  & \beta_i \\
    \beta_j                      & g_{ij}
  \end{matrix}
  \right).
\end{equation}

We integrate each ray until we reach a terminal optical depth of
$\tau_{\rm term}$ at event $x_e^\alpha$.
We choose $\tau_{\rm term}$ large so that
$f_{\rm bound}$ is negligible
(in practice $e^{-\tau_{\rm term}}=0.001$)
and compute $f_{\rm bound}$ assuming radiative equilibrium with the matter,
i.e. $f_{\rm bound}=(1+\exp((\varepsilon-\mu_\nu)/T)$,
and $\mu_{\nu_e}=-\mu_{\bar{\nu}_e}=\mu_{e^{-}}+\mu_p-\mu_n$ and
$\mu_{\nu_x}=0$.
If the ray terminates off the numerical domain
we set $f_{\rm bound}=0$.

Since we don't know the emission event a priori,
\todo{make clear ``emission event'' doesn't involve physical assumption}
we follow the integration backwards in time,
from $t_o$ to $t_e$.
We begin each integration by setting initial values for the variables at $t_o$.
The observer specifies $x^i_o$ and $p_{i,o}$,
and we set $f_{{\rm integ},o}=0$ and $\tau_o=0$.

We integrate Eqns.~\ref{eqn:num_x}-\ref{eqn:num_tau} with adaptive step sizes,
using the 5th-order Dormand-Prince algorithm \cite{pres2007-nr_3rd_ed}.
After a step is taken, the error is estimated for each variable.
If the maximum of these errors exceeds some absolute or relative threshold
the step size is decreased and the step recomputed,
if it is below some threshold the next step size is increased.
% see OdeIntegrators::AdaptiveDense
% and OdeSteppers::DormandPrince5
% and OdeControllers::ProportionalIntegral_Base::StepSizeToAttempt
% and OdeErrorMeasures::ScaledAbsRel
\todo{specify}

The volume data are components of the metric $\psi_{\alpha\beta}$,
its derivatives $\psi_{\alpha\beta,\gamma}$,
and the hydrodynamical data $W$, $u_i$, $\varrho$, $T$, and $Y_e$.
These fields are stored in spectral representation.
If they are computed via hydrodynamical simulation, they are saved to disk
at specified coordinate times and interpolated
with spectral interpolation in space,
and 1st-order polynomial interpolation in time
(as described in \cite[App.~B]{bohn2016-code}).
If computed from a stationary solution to the
general relativistic hydrodynamics equations,
a single time is stored to disk, and no time interpolation is needed.
